---
title: "Cats and humans show correlated affiliative physiological states — Full Analysis Code"
author: "Saho Takagi et al."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(tidyr)
library(brms)
library(purrr)
library(tidybayes)
library(ggplot2)
library(patchwork)
library(loo)
```

# Overview

This script reproduces all statistical analyses and figures reported in
"Cats and humans show correlated affiliative physiological states." It is
organized in the same order as the manuscript:

1. Data import and cleaning
2. Figure 1 / Table S2 — Multivariate model of owner–cat hormonal associations
3. Figure 2 / Table S3 — Hormone–mood associations (primary analysis: individual cat level;
   household-aggregated sensitivity analysis: Table S4)
4. Figure 3 / Table S5 / Table S8 — Proximity, cohabitation duration, mood, and the cat-age
   confound check (primary analysis: individual cat level; household-aggregated sensitivity
   analysis: Table S6)
5. Table S7 — Secure Base Test (SBT) hurdle Gamma models

Mood (POMS2) is reported once per owner and therefore shared across all cats in a household.
Because mood does not vary within a household, including owner identity as a group-level
(random) intercept in cat-level models is non-identifiable (models fail to converge, R-hat > 3).
Dyad-level (cat-level) models without a group-level term are therefore reported as the primary
analysis, with household-level aggregation (mean across cats within a household) retained as a
sensitivity analysis.

Random seeds (`seed = 123`) are fixed throughout for exact reproducibility
of posterior draws, though minor numerical differences may occur across
machines/Stan versions.

# 1. Data Import and Cleaning

```{r load-data}
# Supplementary_rawdata.csv contains 29 cats (rows), one row per cat, with
# 20 columns (ID plus 19 variables). Column names are already clean
# snake_case, so no additional cleaning is required after read.csv().
catdata <- read.csv("Supplementary_rawdata.csv", stringsAsFactors = FALSE)

# Columns (20 total):
# ID, cohabitation_duration_months, cat_age_months, proximity_median_rssi,
# sbt_circle_time_owner_baseline_sec, sbt_circle_time_owner_return_sec,
# sbt_circle_time_stranger_baseline_sec, sbt_circle_time_stranger_return_sec,
# owner_oxytocin, owner_cortisol_creatinine_corrected,
# cat_oxytocin, cat_cortisol_creatinine_corrected,
# poms_anger_hostility, poms_confusion_bewilderment, poms_depression_dejection,
# poms_fatigue_inertia, poms_tension_anxiety, poms_vigor_activity,
# poms_friendliness, poms_total_mood_disturbance
stopifnot(nrow(catdata) == 29)
```

## 1.1 Owner ID mapping

Nineteen owners contributed one to four cats each. This mapping links each
cat (`ID`) to its owner (`owner_id`) and is required for all group-level
(random-effects) models and household-level sensitivity aggregations below.

```{r owner-map}
owner_map <- c(
  C01 = "H01", C02 = "H02", C03 = "H02", C04 = "H03", C05 = "H04",
  C06 = "H05", C07 = "H05", C08 = "H05", C09 = "H05",
  C10 = "H06", C11 = "H07", C12 = "H07", C13 = "H08", C14 = "H09", C15 = "H09",
  C16 = "H10", C17 = "H11", C18 = "H12", C19 = "H12", C20 = "H13", C21 = "H14",
  C22 = "H15", C23 = "H15", C24 = "H16", C25 = "H17", C26 = "H17",
  C27 = "H18", C28 = "H19", C29 = "H19"
)

catdata$owner_id <- owner_map[catdata$ID]
stopifnot(length(unique(catdata$owner_id)) == 19)
```

# 2. Figure 1 / Table S2: Multivariate Model of Owner–Cat Hormonal Associations

Cat oxytocin and cat cortisol concentrations are modeled jointly as correlated
response variables, with owner oxytocin and owner cortisol entered
simultaneously as predictors of both. Owner identity is included as a
group-level (random) intercept for both responses (`(1 | p | owner_id)`,
where `p` links the two response-specific intercepts so their correlation
is estimated).

```{r fig1-data}
df_mv <- catdata %>%
  transmute(
    ID,
    owner_id,
    OT_owner_z   = scale(owner_oxytocin)[, 1],
    Cort_owner_z = scale(owner_cortisol_creatinine_corrected)[, 1],
    OT_cat_z     = scale(cat_oxytocin)[, 1],
    Cort_cat_z   = scale(cat_cortisol_creatinine_corrected)[, 1]
  ) %>%
  filter(!is.na(OT_cat_z), !is.na(Cort_cat_z))

nrow(df_mv)                          # expected: 24
length(unique(df_mv$owner_id))       # expected: 17
```

```{r fig1-model, cache=TRUE}
bf_ot   <- bf(OT_cat_z   ~ OT_owner_z + Cort_owner_z + (1 | p | owner_id))
bf_cort <- bf(Cort_cat_z ~ OT_owner_z + Cort_owner_z + (1 | p | owner_id))

model_mv <- brm(
  bf_ot + bf_cort + set_rescor(TRUE),
  data = df_mv,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1),    class = "b",     resp = "OTcatz"),
    prior(normal(0, 1),    class = "b",     resp = "Cortcatz"),
    prior(exponential(1),  class = "sd",    resp = "OTcatz"),
    prior(exponential(1),  class = "sd",    resp = "Cortcatz"),
    prior(exponential(1),  class = "sigma", resp = "OTcatz"),
    prior(exponential(1),  class = "sigma", resp = "Cortcatz")
  ),
  chains = 4, iter = 6000, warmup = 3000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  cores = 4, seed = 123
)

summary(model_mv)
```

## 2.1 Convergence diagnostics

```{r fig1-diagnostics}
max(rhat(model_mv), na.rm = TRUE)
sum(subset(nuts_params(model_mv), Parameter == "divergent__")$Value)
```

## 2.2 Sensitivity analysis: outlier exclusion

One owner cortisol value was a notable outlier (standardized ≈ 3.5). We
refit the model excluding this observation to confirm the owner
cortisol → cat oxytocin association was not driven by this single point.

```{r fig1-sensitivity, cache=TRUE}
df_mv_excl <- df_mv %>% filter(Cort_owner_z < 3)

model_mv_excl <- update(model_mv, newdata = df_mv_excl)
fixef(model_mv_excl)["OTcatz_Cort_owner_z", ]
```

## 2.3 Correlation matrix (for interpreting the multivariate coefficients)

```{r fig1-cormat}
round(cor(df_mv[, c("OT_owner_z", "Cort_owner_z", "OT_cat_z", "Cort_cat_z")]), 3)
```

## 2.4 Figure 1

```{r fig1-plot, fig.width=10, fig.height=8}
draws <- as_draws_df(model_mv) %>%
  select(
    `Cat OT ~ Owner OT`             = b_OTcatz_OT_owner_z,
    `Cat OT ~ Owner cortisol`       = b_OTcatz_Cort_owner_z,
    `Cat cortisol ~ Owner OT`       = b_Cortcatz_OT_owner_z,
    `Cat cortisol ~ Owner cortisol` = b_Cortcatz_Cort_owner_z
  ) %>%
  tidyr::pivot_longer(everything(), names_to = "path", values_to = "estimate") %>%
  mutate(path = factor(path, levels = rev(c(
    "Cat OT ~ Owner OT",
    "Cat OT ~ Owner cortisol",
    "Cat cortisol ~ Owner OT",
    "Cat cortisol ~ Owner cortisol"
  ))))

p_panelA_fig1 <- draws %>%
  ggplot(aes(x = estimate, y = path)) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
  stat_pointinterval(
    .width = c(0.5, 0.95),
    point_size = 2.3,
    interval_size_range = c(0.8, 2.4)
  ) +
  labs(x = "Standardized posterior estimate", y = NULL) +
  theme_classic(base_size = 16) +
  theme(axis.text.y = element_text(size = 13))

ce_df <- conditional_effects(model_mv, resp = "OTcatz")

p_panelB_fig1 <- ggplot(ce_df[["OTcatz.OTcatz_OT_owner_z"]],
                         aes(x = OT_owner_z, y = estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey80") +
  geom_line(linewidth = 1.2, colour = "black") +
  geom_point(data = df_mv, aes(x = OT_owner_z, y = OT_cat_z),
             inherit.aes = FALSE, size = 3, colour = "grey20") +
  labs(x = "Owner oxytocin (standardized)",
       y = "Cat oxytocin (standardized)") +
  theme_classic(base_size = 16)

p_panelC_fig1 <- ggplot(ce_df[["OTcatz.OTcatz_Cort_owner_z"]],
                         aes(x = Cort_owner_z, y = estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey80") +
  geom_line(linewidth = 1.2, colour = "black") +
  geom_point(data = df_mv, aes(x = Cort_owner_z, y = OT_cat_z),
             inherit.aes = FALSE, size = 3, colour = "grey20") +
  labs(x = "Owner cortisol (standardized)",
       y = NULL) +
  theme_classic(base_size = 16)

fig1 <- free(p_panelA_fig1, side = "l") / (p_panelB_fig1 | p_panelC_fig1) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 16))

fig1
ggsave("Figure1_hormonal_multivariate.png", fig1, width = 10, height = 8, dpi = 400)
```

---

# 3. Figure 2 / Table S3: Hormone–Mood Associations

## 3.1 Cat-level data (primary analysis)

Each cat–owner dyad is modeled individually, without aggregation
(n = 29 dyads for owner-hormone models; n = 24 for cat-hormone models,
reflecting missing cat urine samples). Owner mood is shared across all
cats belonging to the same household; because it does not vary within a
household, a group-level (random) intercept for `owner_id` is
non-identifiable for these models (see Overview) and is therefore omitted.

```{r fig2-data-catlevel}
df_poms_catlevel <- catdata %>%
  select(ID, owner_id, owner_oxytocin, cat_oxytocin,
         owner_cortisol_creatinine_corrected, cat_cortisol_creatinine_corrected,
         proximity_median_rssi, cohabitation_duration_months, cat_age_months,
         starts_with("poms_")) %>%
  mutate(
    OT_owner_z   = scale(owner_oxytocin)[, 1],
    OT_cat_z     = scale(cat_oxytocin)[, 1],
    Cort_owner_z = scale(owner_cortisol_creatinine_corrected)[, 1],
    Cort_cat_z   = scale(cat_cortisol_creatinine_corrected)[, 1],
    proximity_z  = scale(proximity_median_rssi)[, 1],
    breeding_z   = scale(cohabitation_duration_months)[, 1],
    age_z        = scale(cat_age_months)[, 1]
  ) %>%
  pivot_longer(starts_with("poms_"), names_to = "poms", values_to = "score") %>%
  group_by(poms) %>%
  mutate(score_z = scale(score)[, 1]) %>%
  ungroup() %>%
  mutate(
    poms_label = dplyr::recode(poms,
      poms_anger_hostility        = "Anger-Hostility",
      poms_confusion_bewilderment = "Confusion-Bewilderment",
      poms_depression_dejection   = "Depression-Dejection",
      poms_friendliness           = "Friendliness",
      poms_fatigue_inertia        = "Fatigue-Inertia",
      poms_tension_anxiety        = "Tension-Anxiety",
      poms_total_mood_disturbance = "Total mood disturbance",
      poms_vigor_activity         = "Vigor-Activity"
    )
  )

# Expected group sizes:
df_poms_catlevel %>% filter(!is.na(OT_owner_z)) %>% distinct(ID) %>% nrow()  # 29
df_poms_catlevel %>% filter(!is.na(OT_cat_z))   %>% distinct(ID) %>% nrow()  # 24
```

## 3.2 Model-fitting function (cat-level, no group-level term)

```{r fig2-fit-function-catlevel}
fit_poms_catlevel <- function(hormone_var, df) {
  df %>%
    filter(!is.na(.data[[hormone_var]])) %>%
    group_by(poms_label) %>%
    nest() %>%
    mutate(
      model = map(
        data,
        ~ brm(
          formula = as.formula(paste0("score_z ~ ", hormone_var)),
          data = .x,
          family = gaussian(),
          chains = 4, iter = 6000, warmup = 3000,
          control = list(adapt_delta = 0.99, max_treedepth = 15),
          cores = 4, seed = 123
        )
      )
    )
}

check_rhat <- function(model_tbl) {
  model_tbl %>%
    mutate(max_rhat = map_dbl(model, ~ max(rhat(.x), na.rm = TRUE))) %>%
    select(poms_label, max_rhat)
}
```

```{r fig2-fit-models-catlevel, cache=TRUE}
models_OT_owner_catlevel   <- fit_poms_catlevel("OT_owner_z",   df_poms_catlevel)
models_Cort_owner_catlevel <- fit_poms_catlevel("Cort_owner_z", df_poms_catlevel)
models_OT_cat_catlevel     <- fit_poms_catlevel("OT_cat_z",     df_poms_catlevel)
models_Cort_cat_catlevel   <- fit_poms_catlevel("Cort_cat_z",   df_poms_catlevel)

check_rhat(models_OT_owner_catlevel)
check_rhat(models_Cort_owner_catlevel)
check_rhat(models_OT_cat_catlevel)
check_rhat(models_Cort_cat_catlevel)
```

## 3.3 Coefficient extraction, full summary, and Table S3

```{r fig2-extract-catlevel}
extract_poms_coef <- function(model_tbl, hormone_label, coef_name) {
  model_tbl %>%
    mutate(draws = map(model, ~ spread_draws(.x, !!sym(coef_name)))) %>%
    select(poms_label, draws) %>%
    unnest(draws) %>%
    rename(estimate = !!sym(coef_name)) %>%
    group_by(poms_label) %>%
    median_qi(estimate, .width = c(.50, .95)) %>%
    ungroup() %>%
    mutate(hormone = hormone_label)
}

extract_full_summary <- function(model_tbl, hormone_label) {
  model_tbl %>%
    mutate(
      summ = map(model, ~ as.data.frame(summary(.x)$fixed) %>%
                   tibble::rownames_to_column("term"))
    ) %>%
    select(poms_label, summ) %>%
    unnest(summ) %>%
    mutate(hormone = hormone_label)
}

poms_hormone_coef_catlevel <- bind_rows(
  extract_poms_coef(models_OT_owner_catlevel,   "Owner oxytocin", "b_OT_owner_z"),
  extract_poms_coef(models_OT_cat_catlevel,     "Cat oxytocin",   "b_OT_cat_z"),
  extract_poms_coef(models_Cort_owner_catlevel, "Owner cortisol", "b_Cort_owner_z"),
  extract_poms_coef(models_Cort_cat_catlevel,   "Cat cortisol",   "b_Cort_cat_z")
)
print(poms_hormone_coef_catlevel %>% filter(.width == 0.95), n = Inf)

# Full model summaries (Estimate, Est.Error, CI, Rhat, Bulk_ESS, Tail_ESS) for Table S3
s3_catlevel_all <- bind_rows(
  extract_full_summary(models_OT_owner_catlevel,   "Owner oxytocin"),
  extract_full_summary(models_Cort_owner_catlevel, "Owner cortisol"),
  extract_full_summary(models_OT_cat_catlevel,     "Cat oxytocin"),
  extract_full_summary(models_Cort_cat_catlevel,   "Cat cortisol")
)
print(s3_catlevel_all, n = Inf)
write.csv(s3_catlevel_all, "S3_catlevel_raw.csv", row.names = FALSE)
```

## 3.4 Figure 2

```{r fig2-plot-catlevel, fig.width=10, fig.height=8}
p_poms_hormone <- poms_hormone_coef_catlevel %>%
  mutate(
    poms_label = factor(poms_label, levels = rev(c(
      "Total mood disturbance", "Tension-Anxiety", "Depression-Dejection",
      "Confusion-Bewilderment", "Fatigue-Inertia", "Anger-Hostility",
      "Friendliness", "Vigor-Activity"
    ))),
    hormone = factor(hormone, levels = c(
      "Owner oxytocin", "Cat oxytocin", "Owner cortisol", "Cat cortisol"
    ))
  ) %>%
  ggplot(aes(x = estimate, y = poms_label)) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
  geom_linerange(data = ~ filter(.x, .width == 0.95),
                 aes(xmin = .lower, xmax = .upper), linewidth = 0.8) +
  geom_linerange(data = ~ filter(.x, .width == 0.50),
                 aes(xmin = .lower, xmax = .upper), linewidth = 2.4) +
  geom_point(size = 2.3) +
  facet_wrap(~ hormone) +
  labs(x = "Standardized posterior estimate", y = NULL) +
  theme_classic(base_size = 14) +
  theme(
    strip.text = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 13),
    panel.spacing = unit(1.5, "lines")
  )

p_poms_hormone
ggsave("Figure2_hormonal_POMS.png", p_poms_hormone, width = 10, height = 8, dpi = 600)
```

## 3.5 Sensitivity analysis: household-level aggregation (Table S4)

As a sensitivity analysis, all variables are aggregated to the household
level (mean across cats within a household) prior to model fitting.

```{r fig2-data-household-agg}
df_poms_owner_agg <- catdata %>%
  group_by(owner_id) %>%
  summarise(
    owner_oxytocin = mean(owner_oxytocin, na.rm = TRUE),
    owner_cortisol_creatinine_corrected = mean(owner_cortisol_creatinine_corrected, na.rm = TRUE),
    across(starts_with("poms_"), ~ mean(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~ na_if(., NaN))) %>%
  mutate(
    OT_owner_z   = scale(owner_oxytocin)[, 1],
    Cort_owner_z = scale(owner_cortisol_creatinine_corrected)[, 1]
  ) %>%
  pivot_longer(starts_with("poms_"), names_to = "poms", values_to = "score") %>%
  group_by(poms) %>%
  mutate(score_z = scale(score)[, 1]) %>%
  ungroup() %>%
  mutate(
    poms_label = recode(poms,
      poms_anger_hostility        = "Anger-Hostility",
      poms_confusion_bewilderment = "Confusion-Bewilderment",
      poms_depression_dejection   = "Depression-Dejection",
      poms_friendliness           = "Friendliness",
      poms_fatigue_inertia        = "Fatigue-Inertia",
      poms_tension_anxiety        = "Tension-Anxiety",
      poms_total_mood_disturbance = "Total mood disturbance",
      poms_vigor_activity         = "Vigor-Activity"
    )
  )

df_poms_cat_agg <- catdata %>%
  group_by(owner_id) %>%
  summarise(
    cat_oxytocin = mean(cat_oxytocin, na.rm = TRUE),
    cat_cortisol_creatinine_corrected = mean(cat_cortisol_creatinine_corrected, na.rm = TRUE),
    across(starts_with("poms_"), ~ mean(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~ na_if(., NaN))) %>%
  mutate(
    OT_cat_z   = scale(cat_oxytocin)[, 1],
    Cort_cat_z = scale(cat_cortisol_creatinine_corrected)[, 1]
  ) %>%
  pivot_longer(starts_with("poms_"), names_to = "poms", values_to = "score") %>%
  group_by(poms) %>%
  mutate(score_z = scale(score)[, 1]) %>%
  ungroup() %>%
  mutate(
    poms_label = recode(poms,
      poms_anger_hostility        = "Anger-Hostility",
      poms_confusion_bewilderment = "Confusion-Bewilderment",
      poms_depression_dejection   = "Depression-Dejection",
      poms_friendliness           = "Friendliness",
      poms_fatigue_inertia        = "Fatigue-Inertia",
      poms_tension_anxiety        = "Tension-Anxiety",
      poms_total_mood_disturbance = "Total mood disturbance",
      poms_vigor_activity         = "Vigor-Activity"
    )
  )

# Expected group sizes:
n_distinct(df_poms_owner_agg$owner_id)                                         # 19
df_poms_cat_agg %>% filter(!is.na(OT_cat_z)) %>% distinct(owner_id) %>% nrow() # 17
```

```{r fig2-fit-function-household}
fit_poms_by_hormone <- function(hormone_var, df) {
  df %>%
    filter(!is.na(.data[[hormone_var]])) %>%
    group_by(poms_label) %>%
    nest() %>%
    mutate(
      model = map(
        data,
        ~ brm(
          formula = as.formula(paste0("score_z ~ ", hormone_var)),
          data = .x,
          family = gaussian(),
          chains = 4, iter = 6000, warmup = 3000,
          control = list(adapt_delta = 0.99, max_treedepth = 15),
          cores = 4, seed = 123
        )
      )
    )
}
```

```{r fig2-fit-models-household, cache=TRUE}
models_OT_owner_agg   <- fit_poms_by_hormone("OT_owner_z",   df_poms_owner_agg)
models_Cort_owner_agg <- fit_poms_by_hormone("Cort_owner_z", df_poms_owner_agg)
models_OT_cat_agg     <- fit_poms_by_hormone("OT_cat_z",     df_poms_cat_agg)
models_Cort_cat_agg   <- fit_poms_by_hormone("Cort_cat_z",   df_poms_cat_agg)
```

```{r fig2-extract-household}
poms_hormone_coef_agg <- bind_rows(
  extract_poms_coef(models_OT_owner_agg,   "Owner oxytocin", "b_OT_owner_z"),
  extract_poms_coef(models_OT_cat_agg,     "Cat oxytocin",   "b_OT_cat_z"),
  extract_poms_coef(models_Cort_owner_agg, "Owner cortisol", "b_Cort_owner_z"),
  extract_poms_coef(models_Cort_cat_agg,   "Cat cortisol",   "b_Cort_cat_z")
)
print(poms_hormone_coef_agg, n = Inf)

s4_household_all <- bind_rows(
  extract_full_summary(models_OT_owner_agg,   "Owner oxytocin"),
  extract_full_summary(models_Cort_owner_agg, "Owner cortisol"),
  extract_full_summary(models_OT_cat_agg,     "Cat oxytocin"),
  extract_full_summary(models_Cort_cat_agg,   "Cat cortisol")
)
print(s4_household_all, n = Inf)
write.csv(s4_household_all, "S4_household_raw.csv", row.names = FALSE)
```

---

# 4. Figure 3 / Table S5 / Table S8: Proximity, Cohabitation Duration, and Mood

## 4.1 Proximity and cohabitation models (cat-level, primary analysis)

Reuses `df_poms_catlevel` (Section 3.1) and `fit_poms_catlevel()` (Section 3.2).

```{r fig3-catlevel-models, cache=TRUE}
models_proximity_catlevel <- fit_poms_catlevel("proximity_z", df_poms_catlevel)
models_breeding_catlevel  <- fit_poms_catlevel("breeding_z",  df_poms_catlevel)

check_rhat(models_proximity_catlevel)
check_rhat(models_breeding_catlevel)

proximity_coef_catlevel    <- extract_poms_coef(models_proximity_catlevel, "Proximity",     "b_proximity_z")
cohabitation_coef_catlevel <- extract_poms_coef(models_breeding_catlevel,  "Cohabitation", "b_breeding_z")

# Combined coefficient object used in Figure 3 Panel A
catlevel_coef <- bind_rows(
  proximity_coef_catlevel,
  cohabitation_coef_catlevel
)

print(proximity_coef_catlevel, n = Inf)
print(cohabitation_coef_catlevel, n = Inf)

s5_catlevel_all <- bind_rows(
  extract_full_summary(models_proximity_catlevel, "Proximity"),
  extract_full_summary(models_breeding_catlevel,  "Cohabitation")
)
print(s5_catlevel_all, n = Inf)
write.csv(s5_catlevel_all, "S5_catlevel_raw.csv", row.names = FALSE)
```

## 4.2 Cat-age confound check: LOO model comparison (Table S8, primary analysis)

Cohabitation duration and cat age are strongly correlated at the individual
cat level (r = 0.89). To evaluate whether cat age, rather than cohabitation
duration itself, accounts for the associations observed in Section 4.1, we
compare three models (cohabitation alone, cat age alone, both together) for
each POMS2 subscale showing a credible association with cohabitation
duration (Confusion-Bewilderment, Depression-Dejection, Tension-Anxiety,
and Total Mood Disturbance).

```{r fig3-confound-check-catlevel, cache=TRUE}
# Cat-level correlation between cohabitation duration and cat age
cor(catdata$cohabitation_duration_months, catdata$cat_age_months, use = "complete.obs")

fit_confound_check <- function(poms_target, df) {
  d <- df %>% filter(poms == poms_target)

  m_breed <- brm(score_z ~ breeding_z,         data = d, chains = 4, iter = 4000, cores = 4, seed = 123,
                  save_pars = save_pars(all = TRUE))
  m_age   <- brm(score_z ~ age_z,              data = d, chains = 4, iter = 4000, cores = 4, seed = 123,
                  save_pars = save_pars(all = TRUE))
  m_both  <- brm(score_z ~ breeding_z + age_z, data = d, chains = 4, iter = 4000, cores = 4, seed = 123,
                  save_pars = save_pars(all = TRUE))

  list(
    breed = m_breed, age = m_age, both = m_both,
    loo_comparison = loo_compare(
      loo(m_breed, moment_match = TRUE),
      loo(m_age,   moment_match = TRUE),
      loo(m_both,  moment_match = TRUE)
    )
  )
}

result_CB_cat  <- fit_confound_check("poms_confusion_bewilderment", df_poms_catlevel)
result_DD_cat  <- fit_confound_check("poms_depression_dejection",   df_poms_catlevel)
result_TA_cat  <- fit_confound_check("poms_tension_anxiety",        df_poms_catlevel)
result_TMD_cat <- fit_confound_check("poms_total_mood_disturbance", df_poms_catlevel)

result_CB_cat$loo_comparison
result_DD_cat$loo_comparison
result_TA_cat$loo_comparison
result_TMD_cat$loo_comparison

# Convergence check
c(result_CB_cat$breed, result_CB_cat$age, result_CB_cat$both,
  result_DD_cat$breed, result_DD_cat$age, result_DD_cat$both,
  result_TA_cat$breed, result_TA_cat$age, result_TA_cat$both,
  result_TMD_cat$breed, result_TMD_cat$age, result_TMD_cat$both) %>%
  map_dbl(~ max(rhat(.x), na.rm = TRUE))
```

Across all four subscales, the model including cohabitation duration alone
showed the best predictive performance (elpd_diff = 0); adding cat age,
either alone or together with cohabitation duration, did not improve —
and in most cases slightly reduced — predictive performance (see
Supplementary Table S8). This indicates that the observed associations
with cohabitation duration are not attributable to cat age, though the two
variables cannot be fully disentangled given their high collinearity in
this sample (see Limitations).

## 4.3 Figure 3

```{r fig3-plot-catlevel, fig.width=16, fig.height=9}
# ============================================================
# Figure 3
# Panel A: posterior estimates for Proximity / Cohabitation
# Panels B-D: Proximity × VA, DD, TA
# Panels E-G: Cohabitation × VA, DD, TA
# ============================================================

fig3_panelA_data <- catlevel_coef %>%
  filter(hormone %in% c("Proximity", "Cohabitation")) %>%
  mutate(
    poms_label = factor(
      poms_label,
      levels = c(
        "Vigor-Activity",
        "Friendliness",
        "Anger-Hostility",
        "Fatigue-Inertia",
        "Confusion-Bewilderment",
        "Depression-Dejection",
        "Tension-Anxiety",
        "Total mood disturbance"
      )
    )
  )

pd <- position_dodge(width = 0.5)

p_panelA <- ggplot(
  fig3_panelA_data,
  aes(x = estimate, y = poms_label, color = hormone)
) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4, color = "black") +
  geom_linerange(
    data = ~ filter(.x, .width == 0.95),
    aes(xmin = .lower, xmax = .upper),
    linewidth = 0.8,
    position = pd
  ) +
  geom_point(
    data = ~ filter(.x, .width == 0.95),
    size = 2.8,
    position = pd
  ) +
  scale_color_manual(
    values = c("Cohabitation" = "#2CA25F", "Proximity" = "#E6550D"),
    breaks = c("Cohabitation", "Proximity"),
    labels = c("Cohabitation", "Proximity")
  ) +
  labs(x = "Standardized posterior estimate", y = NULL, color = NULL) +
  theme_classic(base_size = 14) +
  theme(legend.position = c(0.93, 0.25), legend.background = element_blank())

make_ce_panel <- function(model, df, poms_key, xvar, xlab, ylab, line_color) {
  d <- df %>% filter(poms == poms_key, !is.na(.data[[xvar]]))
  ce <- conditional_effects(model, effects = xvar)

  ggplot(ce[[xvar]], aes(x = .data[[xvar]], y = estimate__)) +
    geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey80") +
    geom_line(linewidth = 1.2, colour = line_color) +
    geom_point(data = d, aes(x = .data[[xvar]], y = score_z),
               inherit.aes = FALSE, size = 2.5, colour = "grey20") +
    labs(x = xlab, y = ylab) +
    theme_classic(base_size = 13)
}

get_proximity_model <- function(label) {
  models_proximity_catlevel %>% filter(poms_label == label) %>% pull(model) %>% .[[1]]
}

get_breeding_model <- function(label) {
  models_breeding_catlevel %>% filter(poms_label == label) %>% pull(model) %>% .[[1]]
}

p_B1 <- make_ce_panel(get_proximity_model("Vigor-Activity"), df_poms_catlevel,
                       "poms_vigor_activity", "proximity_z",
                       "Proximity (standardized)", "Vigor-Activity (std.)", "#E6550D")
p_B2 <- make_ce_panel(get_proximity_model("Depression-Dejection"), df_poms_catlevel,
                       "poms_depression_dejection", "proximity_z",
                       "Proximity (standardized)", "Depression-Dejection (std.)", "#E6550D")
p_B3 <- make_ce_panel(get_proximity_model("Tension-Anxiety"), df_poms_catlevel,
                       "poms_tension_anxiety", "proximity_z",
                       "Proximity (standardized)", "Tension-Anxiety (std.)", "#E6550D")

p_C1 <- make_ce_panel(get_breeding_model("Vigor-Activity"), df_poms_catlevel,
                       "poms_vigor_activity", "breeding_z",
                       "Cohabitation (standardized)", "Vigor-Activity (std.)", "#2CA25F")
p_C2 <- make_ce_panel(get_breeding_model("Depression-Dejection"), df_poms_catlevel,
                       "poms_depression_dejection", "breeding_z",
                       "Cohabitation (standardized)", "Depression-Dejection (std.)", "#2CA25F")
p_C3 <- make_ce_panel(get_breeding_model("Tension-Anxiety"), df_poms_catlevel,
                       "poms_tension_anxiety", "breeding_z",
                       "Cohabitation (standardized)", "Tension-Anxiety (std.)", "#2CA25F")

panel_B <- p_B1 | p_B2 | p_B3
panel_C <- p_C1 | p_C2 | p_C3

fig3 <-
  free(p_panelA, side = "l") +
  panel_B +
  panel_C +
  plot_layout(
    design = "
    AB
    AC
    ",
    widths = c(2.5, 3),
    heights = c(1, 1)
  ) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 16))

fig3

ggsave("Figure3_proximity_cohabitation_mood.png", fig3,
       width = 16, height = 9, units = "in", dpi = 300)
```

## 4.4 Sensitivity analysis: household-level aggregation (Table S6)

```{r fig3-data-household-agg}
df_cohab_prox <- catdata %>%
  group_by(owner_id) %>%
  summarise(
    breeding_period = mean(cohabitation_duration_months, na.rm = TRUE),
    proximity        = mean(proximity_median_rssi,        na.rm = TRUE),
    across(starts_with("poms_"), ~ mean(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~ na_if(., NaN))) %>%
  mutate(
    breeding_z  = scale(breeding_period)[, 1],
    proximity_z = scale(proximity)[, 1]
  ) %>%
  pivot_longer(starts_with("poms_"), names_to = "poms", values_to = "score") %>%
  group_by(poms) %>%
  mutate(score_z = scale(score)[, 1]) %>%
  ungroup() %>%
  mutate(
    poms_label = recode(poms,
      poms_anger_hostility        = "Anger-Hostility",
      poms_confusion_bewilderment = "Confusion-Bewilderment",
      poms_depression_dejection   = "Depression-Dejection",
      poms_friendliness           = "Friendliness",
      poms_fatigue_inertia        = "Fatigue-Inertia",
      poms_tension_anxiety        = "Tension-Anxiety",
      poms_total_mood_disturbance = "Total mood disturbance",
      poms_vigor_activity         = "Vigor-Activity"
    )
  )
```

```{r fig3-household-models, cache=TRUE}
fit_by_poms_household <- function(poms_target, predictor, df) {
  d <- df %>% filter(poms == poms_target)
  brm(as.formula(paste0("score_z ~ ", predictor)), data = d,
      chains = 4, iter = 6000, warmup = 3000,
      control = list(adapt_delta = 0.99, max_treedepth = 15),
      cores = 4, seed = 123)
}

poms_targets <- c(
  "poms_anger_hostility", "poms_confusion_bewilderment", "poms_depression_dejection",
  "poms_fatigue_inertia", "poms_tension_anxiety", "poms_vigor_activity",
  "poms_friendliness", "poms_total_mood_disturbance"
)
poms_labels_all <- c(
  "Anger-Hostility", "Confusion-Bewilderment", "Depression-Dejection",
  "Fatigue-Inertia", "Tension-Anxiety", "Vigor-Activity",
  "Friendliness", "Total mood disturbance"
)

models_proximity_household <- tibble(poms = poms_targets, poms_label = poms_labels_all) %>%
  mutate(model = map(poms, ~ fit_by_poms_household(.x, "proximity_z", df_cohab_prox)))

models_breeding_household <- tibble(poms = poms_targets, poms_label = poms_labels_all) %>%
  mutate(model = map(poms, ~ fit_by_poms_household(.x, "breeding_z", df_cohab_prox)))

check_rhat(models_proximity_household)
check_rhat(models_breeding_household)

proximity_coef_household    <- extract_poms_coef(models_proximity_household, "Proximity",    "b_proximity_z")
cohabitation_coef_household <- extract_poms_coef(models_breeding_household,  "Cohabitation", "b_breeding_z")

print(proximity_coef_household, n = Inf)
print(cohabitation_coef_household, n = Inf)

s6_household_all <- bind_rows(
  extract_full_summary(models_proximity_household, "Proximity"),
  extract_full_summary(models_breeding_household,  "Cohabitation")
)
print(s6_household_all, n = Inf)
write.csv(s6_household_all, "S6_household_raw.csv", row.names = FALSE)
```

---

# 5. Table S7: Secure Base Test (SBT) Hurdle Gamma Models

## 5.1 Data preparation

```{r sbt-data}
df_in <- catdata %>%
  select(
    ID, owner_id,
    owner_oxytocin, cat_oxytocin,
    owner_cortisol_creatinine_corrected, cat_cortisol_creatinine_corrected,
    starts_with("sbt_circle_time")
  ) %>%
  pivot_longer(
    cols = starts_with("sbt_circle_time"),
    names_to = "condition",
    values_to = "value"
  ) %>%
  mutate(
    condition = sub("^sbt_circle_time_", "", condition),
    condition = sub("_sec$", "", condition)
  ) %>%
  separate(condition, into = c("target", "phase"), sep = "_") %>%
  mutate(
    target = factor(target),
    phase  = factor(phase),
    OT_owner_z   = scale(owner_oxytocin)[, 1],
    OT_cat_z     = scale(cat_oxytocin)[, 1],
    Cort_owner_z = scale(owner_cortisol_creatinine_corrected)[, 1],
    Cort_cat_z   = scale(cat_cortisol_creatinine_corrected)[, 1]
  )

nrow(df_in)                       # 116 (29 cats x 4 owner/stranger x baseline/return; no missing SBT or owner-hormone values)
n_distinct(df_in$owner_id)        # 19 (all owners; brm() below drops the 5 cats with missing cat-hormone values
                                   #     at fit time, leaving an effective n of 24 cats / 17 owners for OT_cat_z / Cort_cat_z)
```

## 5.2 Hormone models

```{r sbt-models, cache=TRUE}
model_in_OTcat <- brm(
  bf(
    value ~ target * phase * OT_cat_z + (1 | owner_id/ID),
    hu    ~ target + phase + OT_cat_z + (1 | owner_id/ID)
  ),
  data = df_in,
  family = hurdle_gamma(),
  chains = 4, iter = 6000, warmup = 3000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  cores = 4, seed = 123
)

max(rhat(model_in_OTcat), na.rm = TRUE)
sum(subset(nuts_params(model_in_OTcat), Parameter == "divergent__")$Value)
summary(model_in_OTcat)

model_in_Cort_cat_z <- brm(
  bf(
    value ~ target * phase * Cort_cat_z + (1 | owner_id/ID),
    hu    ~ target + phase + Cort_cat_z + (1 | owner_id/ID)
  ),
  data = df_in,
  family = hurdle_gamma(),
  chains = 4, iter = 6000, warmup = 3000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  cores = 4, seed = 123
)

max(rhat(model_in_Cort_cat_z), na.rm = TRUE)
sum(subset(nuts_params(model_in_Cort_cat_z), Parameter == "divergent__")$Value)
summary(model_in_Cort_cat_z)
```

---

# 6. Session Information

```{r session-info}
sessionInfo()
```