---
title: "Cats and Humans Share Affiliative Physiological States — Full Analysis Code"
author: "Saho Takagi et al."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(tidyr)
library(brms)
library(purrr)
library(tidybayes)
library(ggplot2)
library(patchwork)
library(loo)
```

# Overview

This script reproduces all statistical analyses and figures reported in
"Cats and Humans Share Affiliative Physiological States." It is organized
in the same order as the manuscript:

1. Data import and cleaning
2. Figure 1 / Table S2 — Multivariate model of owner–cat hormonal associations
3. Figure 2 / Table S3 — Hormone–mood associations (household-aggregated)
4. Figure 3 / Table S4 / Table S6 — Proximity, cohabitation duration, and mood
5. Table S5 — Secure Base Test (SBT) hurdle Gamma models

Random seeds (`seed = 123`) are fixed throughout for exact reproducibility
of posterior draws, though minor numerical differences may occur across
machines/Stan versions.

# 1. Data Import and Cleaning

```{r load-data}
# Supplementary_rawdata.csv contains 29 cats (rows), one row per cat.
# read.csv() converts spaces/parentheses in the original column headers to
# dots (e.g., "breeding period" -> "breeding.period"); we rely on this
# default behavior for consistency with the analyses reported in the paper.
catdata <- read.csv("Supplementary_rawdata.csv", stringsAsFactors = FALSE)

# Expected columns (28 total):
# ID, breeding.period, age.month., beacon.day, beacon.mean, beacon.night,
# SBT.in.owner.Baseline, SBT.in.owner.Return, SBT.in.stranger.Baseline,
# SBT.in.stranger.Return, SBT.follow.owner, SBT.follow.stranger,
# SBT.meow.owner.Baseline, SBT.meow.owner.Alone, SBT.meow.owner.Return,
# SBT.meow.stranger.Baseline, SBT.meow.stranger.Alone, SBT.meow.stranger.Return,
# OT_owner, Ctri_owner, OT_cat, Crti_cat,
# POMS2.AH.ikari., POMS2.CB.konran., POMS2.DD.yokuutsu., POMS2.FI.hirou.,
# POMS2.TA.kincho., POMS2.VA.kakki., POMS2.F.positive., POMS2.TMD
stopifnot(nrow(catdata) == 29)
```

## 1.1 Owner ID mapping

Nineteen owners contributed one to four cats each. This mapping links each
cat (`ID`) to its owner (`owner_id`) and is required for all group-level
(random-effects) models and household-level aggregations below.

```{r owner-map}
owner_map <- c(
  C01 = "H01", C02 = "H02", C03 = "H02", C04 = "H03", C05 = "H04",
  C06 = "H05", C07 = "H05", C08 = "H05", C09 = "H05",
  C10 = "H06", C11 = "H07", C12 = "H07", C13 = "H08", C14 = "H09", C15 = "H09",
  C16 = "H10", C17 = "H11", C18 = "H12", C19 = "H12", C20 = "H13", C21 = "H14",
  C22 = "H15", C23 = "H15", C24 = "H16", C25 = "H17", C26 = "H17",
  C27 = "H18", C28 = "H19", C29 = "H19"
)

catdata$owner_id <- owner_map[catdata$ID]
stopifnot(length(unique(catdata$owner_id)) == 19)
```

# 2. Figure 1 / Table S2: Multivariate Model of Owner–Cat Hormonal Associations

Cat oxytocin and cat cortisol concentrations are modeled jointly as correlated
response variables, with owner oxytocin and owner cortisol entered
simultaneously as predictors of both. Owner identity is included as a
group-level (random) intercept for both responses (`(1 | p | owner_id)`,
where `p` links the two response-specific intercepts so their correlation
is estimated).

```{r fig1-data}
df_mv <- catdata %>%
  transmute(
    ID,
    owner_id,
    OT_owner_z   = scale(OT_owner)[, 1],
    Crti_owner_z = scale(Ctri_owner)[, 1],
    OT_cat_z     = scale(OT_cat)[, 1],
    Crti_cat_z   = scale(Crti_cat)[, 1]
  ) %>%
  filter(!is.na(OT_cat_z), !is.na(Crti_cat_z))

nrow(df_mv)                          # expected: 24
length(unique(df_mv$owner_id))       # expected: 17
```

```{r fig1-model, cache=TRUE}
bf_ot   <- bf(OT_cat_z   ~ OT_owner_z + Crti_owner_z + (1 | p | owner_id))
bf_crti <- bf(Crti_cat_z ~ OT_owner_z + Crti_owner_z + (1 | p | owner_id))

model_mv <- brm(
  bf_ot + bf_crti + set_rescor(TRUE),
  data = df_mv,
  family = gaussian(),
  prior = c(
    prior(normal(0, 1),    class = "b",     resp = "OTcatz"),
    prior(normal(0, 1),    class = "b",     resp = "Crticatz"),
    prior(exponential(1),  class = "sd",    resp = "OTcatz"),
    prior(exponential(1),  class = "sd",    resp = "Crticatz"),
    prior(exponential(1),  class = "sigma", resp = "OTcatz"),
    prior(exponential(1),  class = "sigma", resp = "Crticatz")
  ),
  chains = 4, iter = 6000, warmup = 3000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  cores = 4, seed = 123
)

summary(model_mv)
```

## 2.1 Convergence diagnostics

```{r fig1-diagnostics}
max(rhat(model_mv), na.rm = TRUE)
sum(subset(nuts_params(model_mv), Parameter == "divergent__")$Value)
```

## 2.2 Sensitivity analysis: outlier exclusion

One owner cortisol value was a notable outlier (standardized ≈ 3.5). We
refit the model excluding this observation to confirm the owner
cortisol → cat oxytocin association was not driven by this single point.

```{r fig1-sensitivity, cache=TRUE}
df_mv_excl <- df_mv %>% filter(Crti_owner_z < 3)

model_mv_excl <- update(model_mv, newdata = df_mv_excl)
fixef(model_mv_excl)["OTcatz_Crti_owner_z", ]
```

## 2.3 Correlation matrix (for interpreting the multivariate coefficients)

```{r fig1-cormat}
round(cor(df_mv[, c("OT_owner_z", "Crti_owner_z", "OT_cat_z", "Crti_cat_z")]), 3)
```

---

# 3. Figure 2 / Table S3: Hormone–Mood Associations

We aggregate all variables to the household level (mean across cats within a household) prior to model fitting.

## 3.1 Household-level aggregated data

```{r fig2-data-owner}
df_poms_owner_agg <- catdata %>%
  group_by(owner_id) %>%
  summarise(
    OT_owner   = mean(OT_owner,   na.rm = TRUE),
    Ctri_owner = mean(Ctri_owner, na.rm = TRUE),
    across(starts_with("POMS2"), ~ mean(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~ na_if(., NaN))) %>%
  mutate(
    OT_owner_z   = scale(OT_owner)[, 1],
    Crti_owner_z = scale(Ctri_owner)[, 1]
  ) %>%
  pivot_longer(starts_with("POMS2"), names_to = "poms", values_to = "score") %>%
  group_by(poms) %>%
  mutate(score_z = scale(score)[, 1]) %>%
  ungroup() %>%
  mutate(
    poms_label = recode(poms,
      POMS2.AH.ikari.    = "Anger-Hostility",
      POMS2.CB.konran.   = "Confusion-Bewilderment",
      POMS2.DD.yokuutsu. = "Depression-Dejection",
      POMS2.F.positive.  = "Friendliness",
      POMS2.FI.hirou.    = "Fatigue-Inertia",
      POMS2.TA.kincho.   = "Tension-Anxiety",
      POMS2.TMD          = "Total mood disturbance",
      POMS2.VA.kakki.    = "Vigor-Activity"
    )
  )

df_poms_cat_agg <- catdata %>%
  group_by(owner_id) %>%
  summarise(
    OT_cat   = mean(OT_cat,   na.rm = TRUE),
    Crti_cat = mean(Crti_cat, na.rm = TRUE),
    across(starts_with("POMS2"), ~ mean(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~ na_if(., NaN))) %>%
  mutate(
    OT_cat_z   = scale(OT_cat)[, 1],
    Crti_cat_z = scale(Crti_cat)[, 1]
  ) %>%
  pivot_longer(starts_with("POMS2"), names_to = "poms", values_to = "score") %>%
  group_by(poms) %>%
  mutate(score_z = scale(score)[, 1]) %>%
  ungroup() %>%
  mutate(
    poms_label = recode(poms,
      POMS2.AH.ikari.    = "Anger-Hostility",
      POMS2.CB.konran.   = "Confusion-Bewilderment",
      POMS2.DD.yokuutsu. = "Depression-Dejection",
      POMS2.F.positive.  = "Friendliness",
      POMS2.FI.hirou.    = "Fatigue-Inertia",
      POMS2.TA.kincho.   = "Tension-Anxiety",
      POMS2.TMD          = "Total mood disturbance",
      POMS2.VA.kakki.    = "Vigor-Activity"
    )
  )

# Expected group sizes:
n_distinct(df_poms_owner_agg$owner_id)                                   # 19
df_poms_cat_agg %>% filter(!is.na(OT_cat_z)) %>% distinct(owner_id) %>% nrow()  # 17
```

## 3.2 Model-fitting function (32 univariate models)

```{r fig2-fit-function}
fit_poms_by_hormone <- function(hormone_var, df) {
  df %>%
    filter(!is.na(.data[[hormone_var]])) %>%
    group_by(poms_label) %>%
    nest() %>%
    mutate(
      model = map(
        data,
        ~ brm(
          formula = as.formula(paste0("score_z ~ ", hormone_var)),
          data = .x,
          family = gaussian(),
          chains = 4, iter = 6000, warmup = 3000,
          control = list(adapt_delta = 0.99, max_treedepth = 15),
          cores = 4, seed = 123
        )
      )
    )
}
```

```{r fig2-fit-models, cache=TRUE}
models_OT_owner_agg   <- fit_poms_by_hormone("OT_owner_z",   df_poms_owner_agg)
models_Crti_owner_agg <- fit_poms_by_hormone("Crti_owner_z", df_poms_owner_agg)
models_OT_cat_agg     <- fit_poms_by_hormone("OT_cat_z",     df_poms_cat_agg)
models_Crti_cat_agg   <- fit_poms_by_hormone("Crti_cat_z",   df_poms_cat_agg)
```

## 3.3 Coefficient extraction and Figure 2

```{r fig2-extract}
extract_poms_coef <- function(model_tbl, hormone_label, coef_name) {
  model_tbl %>%
    mutate(draws = map(model, ~ spread_draws(.x, !!sym(coef_name)))) %>%
    select(poms_label, draws) %>%
    unnest(draws) %>%
    rename(estimate = !!sym(coef_name)) %>%
    group_by(poms_label) %>%
    median_qi(estimate, .width = c(.50, .95)) %>%
    ungroup() %>%
    mutate(hormone = hormone_label)
}

poms_hormone_coef <- bind_rows(
  extract_poms_coef(models_OT_owner_agg,   "Owner oxytocin", "b_OT_owner_z"),
  extract_poms_coef(models_OT_cat_agg,     "Cat oxytocin",   "b_OT_cat_z"),
  extract_poms_coef(models_Crti_owner_agg, "Owner cortisol", "b_Crti_owner_z"),
  extract_poms_coef(models_Crti_cat_agg,   "Cat cortisol",   "b_Crti_cat_z")
)

print(poms_hormone_coef, n = Inf)
```

```{r fig2-full-summary}
# Full model summaries (Rhat, Bulk_ESS, Tail_ESS) for Table S3
extract_full_summary <- function(model_tbl, hormone_label) {
  model_tbl %>%
    mutate(
      summ = map(model, ~ as.data.frame(summary(.x)$fixed) %>%
                   tibble::rownames_to_column("term"))
    ) %>%
    select(poms_label, summ) %>%
    unnest(summ) %>%
    mutate(hormone = hormone_label)
}

s3_all <- bind_rows(
  extract_full_summary(models_OT_owner_agg,   "Owner oxytocin"),
  extract_full_summary(models_Crti_owner_agg, "Owner cortisol"),
  extract_full_summary(models_OT_cat_agg,     "Cat oxytocin"),
  extract_full_summary(models_Crti_cat_agg,   "Cat cortisol")
)
print(s3_all, n = Inf)
```

```{r fig2-plot, fig.width=10, fig.height=8}
p_poms_hormone <- poms_hormone_coef %>%
  mutate(
    poms_label = factor(poms_label, levels = rev(c(
      "Total mood disturbance", "Tension-Anxiety", "Depression-Dejection",
      "Confusion-Bewilderment", "Fatigue-Inertia", "Anger-Hostility",
      "Friendliness", "Vigor-Activity"
    ))),
    hormone = factor(hormone, levels = c(
      "Owner oxytocin", "Cat oxytocin", "Owner cortisol", "Cat cortisol"
    ))
  ) %>%
  ggplot(aes(x = estimate, y = poms_label)) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4) +
  geom_linerange(data = ~ filter(.x, .width == 0.95),
                 aes(xmin = .lower, xmax = .upper), linewidth = 0.8) +
  geom_linerange(data = ~ filter(.x, .width == 0.50),
                 aes(xmin = .lower, xmax = .upper), linewidth = 2.4) +
  geom_point(size = 2.3) +
  facet_wrap(~ hormone) +
  labs(x = "Standardized posterior estimate", y = NULL) +
  theme_classic(base_size = 14) +
  theme(
    strip.text = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 13),
    panel.spacing = unit(1.5, "lines")
  )

p_poms_hormone
ggsave("Figure2_hormonal_POMS.png", p_poms_hormone, width = 10, height = 8, dpi = 600)
```

---

# 4. Figure 3 / Table S4 / Table S6: Proximity, Cohabitation Duration, and Mood

## 4.1 Household-level aggregated data (proximity, cohabitation duration, cat age)

```{r fig3-data}
df_cohab_prox <- catdata %>%
  group_by(owner_id) %>%
  summarise(
    breeding_period = mean(breeding.period, na.rm = TRUE),
    cat_age         = mean(age.month.,      na.rm = TRUE),
    proximity       = mean(beacon.mean,     na.rm = TRUE),
    across(starts_with("POMS2"), ~ mean(.x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~ na_if(., NaN))) %>%
  mutate(
    breeding_z  = scale(breeding_period)[, 1],
    age_z       = scale(cat_age)[, 1],
    proximity_z = scale(proximity)[, 1]
  ) %>%
  pivot_longer(starts_with("POMS2"), names_to = "poms", values_to = "score") %>%
  group_by(poms) %>%
  mutate(score_z = scale(score)[, 1]) %>%
  ungroup() %>%
  mutate(
    poms_label = recode(poms,
      POMS2.AH.ikari.    = "Anger-Hostility",
      POMS2.CB.konran.   = "Confusion-Bewilderment",
      POMS2.DD.yokuutsu. = "Depression-Dejection",
      POMS2.F.positive.  = "Friendliness",
      POMS2.FI.hirou.    = "Fatigue-Inertia",
      POMS2.TA.kincho.   = "Tension-Anxiety",
      POMS2.TMD          = "Total mood disturbance",
      POMS2.VA.kakki.    = "Vigor-Activity"
    )
  )

# Confounding check: cohabitation duration vs. cat age (household level)
cor(df_cohab_prox$breeding_z, df_cohab_prox$age_z, use = "complete.obs")
```

## 4.2 Proximity models (8 POMS subscales)

```{r fig3-proximity-models, cache=TRUE}
fit_proximity_by_poms <- function(poms_target, df) {
  d <- df %>% filter(poms == poms_target)
  brm(score_z ~ proximity_z, data = d,
      chains = 4, iter = 6000, warmup = 3000,
      control = list(adapt_delta = 0.99, max_treedepth = 15),
      cores = 4, seed = 123)
}

poms_targets <- c(
  "POMS2.AH.ikari.", "POMS2.CB.konran.", "POMS2.DD.yokuutsu.",
  "POMS2.FI.hirou.", "POMS2.TA.kincho.", "POMS2.VA.kakki.",
  "POMS2.F.positive.", "POMS2.TMD"
)

models_proximity <- tibble(poms = poms_targets) %>%
  mutate(
    poms_label = recode(poms,
      POMS2.AH.ikari.    = "Anger-Hostility",
      POMS2.CB.konran.   = "Confusion-Bewilderment",
      POMS2.DD.yokuutsu. = "Depression-Dejection",
      POMS2.F.positive.  = "Friendliness",
      POMS2.FI.hirou.    = "Fatigue-Inertia",
      POMS2.TA.kincho.   = "Tension-Anxiety",
      POMS2.TMD          = "Total mood disturbance",
      POMS2.VA.kakki.    = "Vigor-Activity"
    ),
    model = map(poms, ~ fit_proximity_by_poms(.x, df_cohab_prox))
  )

models_proximity %>%
  mutate(max_rhat = map_dbl(model, ~ max(rhat(.x), na.rm = TRUE))) %>%
  select(poms_label, max_rhat)

proximity_coef <- extract_poms_coef(models_proximity, "Proximity (beacon RSSI)", "b_proximity_z")
print(proximity_coef, n = Inf)
```

## 4.3 Cohabitation duration models (8 POMS subscales)

```{r fig3-cohabitation-models, cache=TRUE}
fit_breeding_by_poms <- function(poms_target, df) {
  d <- df %>% filter(poms == poms_target)
  brm(score_z ~ breeding_z, data = d,
      chains = 4, iter = 4000, warmup = 2000, cores = 4, seed = 123)
}

remaining_poms <- c("POMS2.AH.ikari.", "POMS2.FI.hirou.", "POMS2.VA.kakki.",
                     "POMS2.F.positive.", "POMS2.TMD")

models_breeding_remaining <- tibble(poms = remaining_poms) %>%
  mutate(
    poms_label = recode(poms,
      POMS2.AH.ikari.   = "Anger-Hostility",
      POMS2.FI.hirou.   = "Fatigue-Inertia",
      POMS2.VA.kakki.   = "Vigor-Activity",
      POMS2.F.positive. = "Friendliness",
      POMS2.TMD         = "Total mood disturbance"
    ),
    model = map(poms, ~ fit_breeding_by_poms(.x, df_cohab_prox))
  )
```

## 4.4 LOO model comparison: cohabitation duration vs. cat age (Table S6)

Because cohabitation duration and cat age are strongly correlated even
after household-level aggregation, we compare three models (cohabitation
alone, cat age alone, both together) for each POMS2 subscale showing an
association with cohabitation duration in preliminary analyses
(Depression-Dejection, Tension-Anxiety, Confusion-Bewilderment).

```{r fig3-confound-check, cache=TRUE}
fit_confound_check <- function(poms_target, df) {
  d <- df %>% filter(poms == poms_target)

  m_breed <- brm(score_z ~ breeding_z,             data = d, chains = 4, iter = 4000, cores = 4, seed = 123)
  m_age   <- brm(score_z ~ age_z,                  data = d, chains = 4, iter = 4000, cores = 4, seed = 123)
  m_both  <- brm(score_z ~ breeding_z + age_z,      data = d, chains = 4, iter = 4000, cores = 4, seed = 123)

  list(
    breed = m_breed, age = m_age, both = m_both,
    loo_comparison = loo_compare(
      loo(m_breed, moment_match = TRUE),
      loo(m_age,   moment_match = TRUE),
      loo(m_both,  moment_match = TRUE)
    )
  )
}

result_DD <- fit_confound_check("POMS2.DD.yokuutsu.", df_cohab_prox)
result_TA <- fit_confound_check("POMS2.TA.kincho.",   df_cohab_prox)
result_CB <- fit_confound_check("POMS2.CB.konran.",   df_cohab_prox)

result_DD$loo_comparison
result_TA$loo_comparison
result_CB$loo_comparison

# Reuse the already-fitted "cohabitation alone" models for the remaining
# 5 subscales (AH, FI, VA, F, TMD) together with DD/TA/CB above.
models_breeding_existing <- tibble(
  poms_label = c("Depression-Dejection", "Tension-Anxiety", "Confusion-Bewilderment"),
  model = list(result_DD$breed, result_TA$breed, result_CB$breed)
)

models_breeding_all8 <- bind_rows(models_breeding_existing, models_breeding_remaining)

cohabitation_coef <- extract_poms_coef(models_breeding_all8, "Cohabitation", "b_breeding_z")
print(cohabitation_coef, n = Inf)
```

## 4.5 Figure 3

```{r fig3-plot, fig.width=14, fig.height=8}
fig3_panelA_data <- bind_rows(cohabitation_coef, proximity_coef) %>%
  mutate(
    poms_label = factor(poms_label, levels = rev(c(
      "Total mood disturbance", "Tension-Anxiety", "Depression-Dejection",
      "Confusion-Bewilderment", "Fatigue-Inertia", "Anger-Hostility",
      "Friendliness", "Vigor-Activity"
    )))
  )

p_panelA <- fig3_panelA_data %>%
  ggplot(aes(x = estimate, y = poms_label, color = hormone)) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.4, color = "black") +
  geom_linerange(data = ~ filter(.x, .width == 0.95),
                 aes(xmin = .lower, xmax = .upper),
                 linewidth = 0.8, position = position_dodge(width = 0.5)) +
  geom_linerange(data = ~ filter(.x, .width == 0.50),
                 aes(xmin = .lower, xmax = .upper),
                 linewidth = 2.4, position = position_dodge(width = 0.5)) +
  geom_point(size = 2.3, position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("Cohabitation" = "#2CA25F", "Proximity (beacon RSSI)" = "#E6550D"),
                      labels = c("Cohabitation", "Proximity")) +
  labs(x = "Standardized posterior estimate", y = NULL, color = NULL) +
  theme_classic(base_size = 14) +
  theme(legend.position = c(0.85, 0.15))

m_proximity_va <- models_proximity %>%
  filter(poms_label == "Vigor-Activity") %>%
  pull(model) %>% .[[1]]

d_va <- df_cohab_prox %>% filter(poms == "POMS2.VA.kakki.")

ce_va <- conditional_effects(m_proximity_va, effects = "proximity_z")

p_panelB <- ggplot(ce_va[["proximity_z"]],
                    aes(x = proximity_z, y = estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey80") +
  geom_line(linewidth = 1.2, colour = "#E6550D") +
  geom_point(data = d_va, aes(x = proximity_z, y = score_z),
             inherit.aes = FALSE, size = 3, colour = "grey20") +
  labs(x = "Proximity (standardized beacon RSSI)",
       y = "Vigor-Activity (standardized)") +
  theme_classic(base_size = 16)

fig3 <- free(p_panelA, side = "l") | p_panelB
fig3 <- fig3 +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 19))

fig3
ggsave("Figure3_proximity_cohabitation_mood.png", fig3, width = 14, height = 8, dpi = 400)
```

---

# 5. Table S5: Secure Base Test (SBT) Hurdle Gamma Models

## 5.1 Data preparation

```{r sbt-data}
df_in <- catdata %>%
  select(
    ID, owner_id,
    OT_owner, OT_cat, Ctri_owner, Crti_cat,
    starts_with("SBT.in")
  ) %>%
  pivot_longer(
    cols = starts_with("SBT.in"),
    names_to = "condition",
    values_to = "value"
  ) %>%
  separate(condition, into = c("SBT", "behavior", "target", "phase"), sep = "\\.") %>%
  mutate(
    target = factor(target),
    phase  = factor(phase),
    OT_owner_z   = scale(OT_owner)[, 1],
    OT_cat_z     = scale(OT_cat)[, 1],
    Crti_owner_z = scale(Ctri_owner)[, 1],
    Crti_cat_z   = scale(Crti_cat)[, 1]
  )

nrow(df_in)                       # expected: 96 (29 cats x ~4 owner/stranger x baseline/return, minus any NA)
n_distinct(df_in$owner_id)        # expected: 17 (cats with valid hormone data)
```

## 5.2 Hormone models

```{r sbt-models, cache=TRUE}
model_in_OTcat <- brm(
  bf(
    value ~ target * phase * OT_cat_z + (1 | owner_id/ID),
    hu    ~ target + phase + OT_cat_z + (1 | owner_id/ID)
  ),
  data = df_in,
  family = hurdle_gamma(),
  chains = 4, iter = 6000, warmup = 3000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  cores = 4, seed = 123
)

max(rhat(model_in_OTcat), na.rm = TRUE)
sum(subset(nuts_params(model_in_OTcat), Parameter == "divergent__")$Value)
summary(model_in_OTcat)

model_in_Crti_cat_z <- brm(
  bf(
    value ~ target * phase * Crti_cat_z + (1 | owner_id/ID),
    hu    ~ target + phase + Crti_cat_z + (1 | owner_id/ID)
  ),
  data = df_in,
  family = hurdle_gamma(),
  chains = 4, iter = 6000, warmup = 3000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  cores = 4, seed = 123
)

max(rhat(model_in_Crti_cat_z), na.rm = TRUE)
sum(subset(nuts_params(model_in_Crti_cat_z), Parameter == "divergent__")$Value)
summary(model_in_Crti_cat_z)
```

---

# 6. Session Information

```{r session-info}
sessionInfo()
```
